---
title: "Homework "
author: "Eric Gleiter; Git hub: Eric-Gleiter"
date: "2/19/2020"
output: html_document
---
*Note: All package checks and installations code is hidden on the html* 

```{r echo=FALSE, results='hide', message=FALSE}
if (!require("ggplot2")) {
install.packages("ggplot2")
}

if (!require("dplyr")) {
install.packages("dplyr")
}

library(ggplot2)
library(dplyr)
library(datasets)


```


**Question 1)**
```{r}
day0 <- ChickWeight %>% filter(Time == 0)
weight0 <- day0 %>%
  group_by(Diet) %>%
  summarize( n= n(),
            average_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight,  na.rm = TRUE)
)

weight0
```



**Question 2)**
```{r}
check_complete <- ChickWeight %>%
  group_by(Chick) %>%
  tally() %>%
  filter(n!=12)

#gives that chicks, 18, 16, 15, 8, and 44 do not have complete weight data

complete <- ChickWeight %>%
  filter(Chick != 18 &
           Chick != 16 &
           Chick != 15 &
           Chick != 8 &
           Chick != 44)

head(complete,3)
```


**Question 3)**
```{r}
complete <- complete %>%
  group_by(Chick) %>%
  mutate(start_weight = first(weight, order_by = Chick)) %>%
  mutate(weightgain = weight - start_weight)

head(complete,3)
```


**Question 4)**
```{r}
day21 <- complete %>%
  filter(Time==21)

ggplot(data=day21, aes(x= reorder(Diet,weightgain,FUN=median),y=weightgain)) + geom_boxplot() + xlab("Diet") + ylab("Weight Gain on Day 21") + labs(title = "Weight Gain vs Diet for day 21") + ylim(0,350)
```
The relationship between medain weight and diet is clear after 21 days chicks on diet 3 are the heaviest followed by deit 4 which are in turn follwed by chicks on diet 2 and the lightest chicks are those on diet 1. It is interesting to note that while diet 4 does not have a higher median weight than diet 3 is it less spread out and may be considered superior depending on the application.


**Question 5)**
```{r}
ggplot(day21 , aes(x=Chick, y=weightgain)) + geom_bar(stat="identity") + ylab("Weight Gain on Day 21") + labs(title = "Weight Gain on day 21 for each Chick")
maxChick <- day21 %>% filter(weightgain == max(weightgain))
minChick <- day21 %>% filter(weightgain == min(weightgain))

head(maxChick)
head(minChick)
```

As the graph shows (Which is order by deit and increasing order) Chick 24 on diet 2 has the least weight gain with only a gain of 32 while chick 35 on diet 3 had the largest weightgain with a gain of 332. these Results are supported by the maxChick and minChick dataframe with show the max and min weightgain for each diet.


**Question 6)**
```{r}

```



